Stack Safety as a Security Property

### 2020-07-09 main.tex

To phrase these two aspects of stack safety in terms of security
properties \leo{in terms of security properties or ``using security
  language''?}  requires an explicit attacker model: a clear
understanding of who is being protected and from what. Most existing
stack protection mechanisms deal with scenarios where a callee gets
influenced by external inputs and ends up executing potentially
malicious code. To avoid the details of exactly how this happens and
keep the presentation clear, we pick an even stronger model: when
stating the protections that a function can rely on when it makes a
call, we assume that the callee can execute arbitrary instructions
until it returns.

Adopting that model in our setting, however, poses another novel
challege: not only does a caller need to be protected from its callee,
but the callee itself needs to be protected from {\em its}
callees. That is, the notion of ``attacker'' changes throughout the
execution of a program!  We address this problem by using a similarly
changing notion of accessibility for the different components of the
machine: we adjust the integrity and confidentiality level of
different stack locations and registers at every transition (call or
return) between functions.  \bcp{that's way too terse!}\leo{Slightly
  better, still needs a bit probably}
%\apt{connection to ``attacker'' still vague.}  \rb{Mention
%  to varying ``accessing power'' of the attacker?  The notion of
%  attacker changes on certain actions that are not externally
%  visible.} \bcp{Maybe the best way is to grab the bull by the horns:
%  ``If we're going to phrase stack safety as a security property, we
%  need to talk about an attacker model, and this is where the
%  technicalities get really interesting...''}

@@

%% <<<<<<< HEAD
%% \definition
%% %  Let $\callmap$ a call map and $\codemap$ a function map.
%%   Given a state $(\mach, \pol) \in \MPS$ that is well-formed w.r.t. a
%%   variant stack $\vs$, we say that an MP-trace induced from $(\mach,
%%   \pol)$ satisfies the \emph{strong stack safety} at a given call
%%   depth as follows.
%% % Refactor repetitive notation before case split
%%   In all cases that let $(\mach, \pol) \hookrightarrow_P
%%   (\mach', \pol', \obs)$.
%%   %and assume (unless the step is a return) that the
%%   %variant stack can step, $\vs \hookrightarrow \vs'$, i.e., all its variant
%%   %machines can step.

>>>>

To phrase these two aspects of stack safety as security properties requires
an explicit attacker model: a clear understanding of who is being protected
and from what. Most existing stack protection mechanisms envisage scenarios
where a callee is influenced by external inputs to corrupt the stack in some
way. To avoid getting into the details of exactly how this happens when
formalizing our properties, we use an even stronger model: when we state the
protections that a function can rely on when it makes a call, we assume that
the callee can execute arbitrary instructions until it returns.

A key challenge in adapting the concepts of language-based security to this
setting is that stack safety is a ``nested'' property: not only does a given
caller need to be protected from its callee, but the callee also needs to be
protected from {\em its} callees, and so on. That is, the notion of
``attacker'' changes throughout the execution of the program.  We address
this issue using a dynamically changing notion of {\em accessibility} for
the different components of the machine state and adjusting the integrity
and confidentiality levels of different stack locations and registers at
every transition (call or return) between functions.

@@

%% \definition
%% %  Let $\callmap$ a call map and $\codemap$ a function map.
%%   Given a state $(\mach, \pol) \in \MPS$ that is well-formed w.r.t. a
%%   variant stack $\vs$, we say that an MP-trace induced from $(\mach,
%%   \pol)$ satisfies the \emph{strong stack safety} at a given call
%%   depth as follows.
%% % Refactor repetitive notation before case split
%%   In all cases that let $(\mach, \pol) \hookrightarrow_P
%%   (\mach', \pol', \obs)$.
%%   %and assume (unless the step is a return) that the
%%   %variant stack can step, $\vs \hookrightarrow \vs'$, i.e., all its variant
%%   %machines can step.
